Stack Safety as a Security Property

### 2020-07-09 main.tex
Work on intro

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

@@

To phrase these two aspects of stack safety in terms of security
properties \leo{in terms of security properties or ``using security
  language''?}  requires an explicit attacker model: an clear
understanding of who is being protected and from what. Most existing
stack protection mechanisms deal with scenarios where a callee gets
influenced by external inputs and ends up executing potentially
malicious code. To avoid the details of exactly how this happens and
keep the presentation clear, we pick an even stronger model: when
stating the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns.

Adopting that model in our setting, however, poses another novel
challege: not only does a caller need to be protected from its callee,
but the callee itself needs to be protected from {\em its}
callees. That is, the notion of ``attacker'' changes throughout the
execution of a program!  We address this problem by using a similarly
changing notion of accessibility for the different components of the
machine: we adjust the integrity and confidentiality level of
different stack locations and registers at every transition (call or
return) between functions.  \bcp{that's way too terse!}\leo{Slightly
  better, still needs a bit probably}
%\apt{connection to ``attacker'' still vague.}  \rb{Mention
%  to varying ``accessing power'' of the attacker?  The notion of
%  attacker changes on certain actions that are not externally
%  visible.} \bcp{Maybe the best way is to grab the bull by the horns:
%  ``If we're going to phrase stack safety as a security property, we
%  need to talk about an attacker model, and this is where the
%  technicalities get really interesting...''}

@@

\section{Threat Model}
\label{sec:threat}

@@

\apt{Informal definition of policies as auxiliary state here.}

@@

In this section we describe our machine model. Our definition of stack
safety is generic and this section discusses an abstract interface to
our machine model and some kind of static or dynamic, software- or
hardware-enforced security monitor. We will refer back to the running
example of the previous section for concreteness, and then,
in~\cref{sec:enforcement}, extend this model with a dynamic policy
enforcement mechanism based on
\citet{DBLP:conf/sp/RoesslerD18}.\apt{needs a relook}

@@

The basic building blocks of the machine are {\em values} and {\em
  addresses}. Both are drawn from some set of {\em words}
$\WORDS$, ranged over by $\word$.
%
Our machine states are composed of {\em components} $k$, addressed either by
words (memory addresses) or by register names ($r$) drawn from some set
$\REGS$.
%
    \[\component \in \COMPONENTS = \WORDS + \REGS \]
%
The register names are assumed to include two special purpose registers: the
program counter {\PCname} and the stack pointer \SP.

@@

The step function of the machine takes a machine state and produces a
pair of a machine state and an {\em observation}, which may be either
silent (written $\tau$) or a word of data that the machine
communicates to the outside world:
%
\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
\[\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \]
%

Notice that we choose to model observations explicitly, rather than simply treating
a subset of registers or memory as being observable.  This choice gives
us the flexibility to treat programs as observationally equivalent even when
their internal states may differ.
%The choice of modelling observations is an interesting one. A first approach
%would be to consider any change to the registers or memory as observable.
%However, as some parts of the stack will be considered ``private''
%this choice won't work. A relaxation of this to consider all changes
%to ``non-private'' data as observable could work in principle, but
%does not allow for an interesting weaker-but-more-efficient enforcement
%mechanism. We will come back to this point in \Cref{sec:lazy}.

In the running example of \Cref{sec:running-example}, 
$\WORDS$ is the
set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
\cup \{\PCname,\SP\}$. Observations are the writes to the
  special $\rout$ location, which can be thought of as a memory-mapped
  output port: if during an execution step a word $w$ is
written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.

\subsection{Policies}

A policy is a general model of an enforcement mechanism, consisting of a
set of policy states \(\pol \in \POLS\), a policy step function
\((\mach, \pol) \polstep \pol' : \MACHS \times \POLS
\rightharpoonup \POLS\), and an initializer \(\pinit : \MACHS \rightharpoonup
\POLS\). The initializer establishes the initial policy state for a given
machine state, but is partial to reflect that some policies are only defined
on states that are well-formed in some way.\apt{Is this our current thinking?}

We write $\MPS$ for the set of all pairs of machine states and policy
states.
%
\[\mpstatename \in \MPS = \MACHS \times \POLS\]
%
We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.

\apt{TODO: relate back to \Cref{sec:threat} and forward to \Cref{sec:enforcement}.}
%\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
%in more detail that the intention is to capture any kind of dynamic
%enforcement mechanism and give as many names and citations as we can.}

@@

In our model, the machine step function is total -- a program does not halt
on its own, but may instead enter a silent terminal loop. Runs are only cut
short by the policy step function. This means that software-only policies that
can catch errors must be modeled as reaching an error state\apt{??}, not 
halting outright.\apt{Is this our current thinking?}

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by the step function on $\MPS$. Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final observation is
a \(\tau\):

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \mpstate{\mach_0}{\pol_0} \polstep \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstopol{\obs}
                 \mpstate{\mach_1}{\pol_1}\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow
                 \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
\end{minipage}

@@

\paragraph*{Until}
The operation \(\PUT~(\Trace,f)\)
%\bcp{Wouldn't
%``While'' be more mnemonic?} \sna{Appealing, but it's more of a
%``while-not''? Trying out ``until''. In any case... moving away
%from doing joins means this needs to get some work to eliminate
%the trailing observation. Here's an attempt.}
takes a trace
\(\Trace : \mathit{trace} ~ \TRACEELEMS \times \OBSS\) and a predicate
on elements \(f \subset \TRACEELEMS\) and gives the prefix of
\(\Trace\) ending with the first element on which \(f\) holds.
Subtly, because the observation associated with a trace represents its
next step, when we take a prefix from a trace containing observations
we must replace the final observation with a \(\tau\).

  \begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f\ \traceelem\)}
             {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
  \judgment{}
           {\(\PUT~(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT~(f,\Trace) = \Trace'\)}
              {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
  \end{minipage}
  \end{center}
%
Note that if the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds,
\(\PUT\) is the identity function.

@@

\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
            {\(\PUT~(f,\machT')
              = \machT\)}
            {\(\mach \hookrightarrow \machT | f\)}
\noindent
And similarly for traces with policies:

\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
            {\(\PUT~(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
         {\(\mpstatename \hookrightarrow \MPT | f\)}

@@

We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
as far as an external observer is concerned, if the sequence of
non-silent observations of $\obsT_1$ is a prefix of those of
$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
coinductively:

@@

\leo{I still don't know how to do this transition. Maybe contrast
with the ``perhaps surprisingly'' later?} Note that
similarity of observation traces could alternatively be defined coinductively,
just like the prefix relation, but by dropping the asymmetric rules that allow
for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}

Note that an infinite silent trace is a
prefix of (and equivalent to) any other trace. While this might seem
surprising at first, it makes sense in a timing-insensitive context:
an external observer looking at two machine runs cannot (computably)
distinguish between a machine that steps forever and a machine that
steps for a long time before producing some output.

@@

In this section we formalize our definition of stack safety.We
begin with our model of \leo{compiler-provided?} knowledge about the
program structure, and then dive into the formal definitions of local state
encapsulation and well-bracketed control flow that together constitute
stack safety.

@@

\paragraph*{Return Map}
Similarly, a {\em return map} identifies addresses that complete a
return sequence. Formally, it is simply a set of words; each one
is the address of the final jump corresponding to a function return:
%
\[\retmap \in \RETMAPS \subseteq \WORDS\]
%
In our running example, the return map identifies the $\jalr$
instructions 36 and 47:
\[ \retmap = \{36, 47\} \]

@@

We separately identify, for any call state, what it means for a state
to have correctly returned from a call. We consider the call to have
returned when control returns to the instruction following
the \jal with the stack restored. Formally, this is a predicate on
states that says when the machine has just returned from a specified call state
(where $\nextPC$ gives the address of the next instruction after a given address):
\[%\begin{split}
    \ret{\mach_c} \triangleq  \lambda \mach_r . \mach_r(\PCname) =
    \nextPC(\mach_c(\PCname)) \land  \mach_r(\SP) = \mach_c(\SP)
%\end{split}
\]
There are two such return points in our running example, both of which
are depicted in the figure: the state before the execution of
instruction 27 corresponding to the return from {\tt h} and the state
before instruction 8 corresponding to the return from {\tt g}.

@@

In the rest of this section we implictly parameterize all 
definitions by a function map $\codemap$, a call map $\callmap$, a return
map $\retmap$ and an entry map $\entmap$.

@@

When \(\mach_c\) is a call state (\(\callmap(\mach_c(\PCname)) = n\) for some $n$) and
\(\mach_c \stepstoobs{\obs} \mach_e\) for some 
$\mach_e$ and $\obs$, we can construct a contour based on the bounds of the
callee's stack frame.
%
The private state of the callee is the entire area of the memory from
stack base to stack pointer, with the exception of the $n$ locations
corresponding to the arguments and result locations of the
call. Private components are marked high confidentiality and high
integrity: the callee should be able to neither read nor write them.
%
Everything above the stack pointer at the point of the call is
``uninitialized'' memory that can be written into but should not be
read. To protect the callee itself from such reads, these components
are marked high confidentiality and low integrity.
%
The same is true for the designated output location: it can be
freely written into but should not be read.
%
Code is always high integrity and low confidentiality.
%
Finally, registers and the frame of arguments and results are free
for the callee to use, and are therefore tagged with low integrity and
low confidentiality.

 \[\mathit{Cof}(\mach,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and } \codemap(\component) = \funid \text{ for some }\funid\\
    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \geq \mach(\SP)
                          \text{ or if } \component = \rout\\
    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \leq \mach(\SP) \negate  n \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]

Going back to our running example, here is the state of the machine at
the call from {\tt main} to {\tt g}, showing addresses annotated with their 
corresponding labels (not shown: the labels of the instructions
corresponding to {\tt g} that are $(\LOWSEC, \HIGHINT)$ instead of
$(\HIGHSEC, \HIGHINT)$)\apt{see previous note}:
\vspace*{0.2em}
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{$(\HIGHSEC,\HIGHINT)$}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\emptyoutc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{$(\HIGHSEC, \HIGHINT)$}}]%
\memory{2}{\fcallc}[{\makebox[0pt]{$(\LOWSEC, \LOWINT)$}}]%
\memory{6}{\unusedc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]
~$\cdots$
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\MemoryLabel{-38.5em}{0.75em}{0}
\MemoryLabel{-34.5em}{0.75em}{0}
\\
\end{center}

@@

\subsubsection{Stack Integrity}

@@

Let \(\contour\) be a contour and \(\MPT\) be a (potentially infinite)
\(\MPOTRACE\). We will say \(\MPT\) satisfies {\em trace integrity} with
respect to \(\contour\) if its final state (if any) agrees with its
initial one on all components marked \(\HIGHSEC\) in \(\contour\):
%
\[\begin{split}
  \propdef{TraceIntegrity} ~ \contour ~ \MPT \triangleq &
    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
    \mach_e(\component) = \mach_r(\component) \\
\end{split}\]
\sna{Experimenting with font on the mathy version of the property.}

Intuitively, the caller expects to find their private data untouched
after the callee returns. That means that if the callee never returns,
integrity is trivially satisfied. Similarly, the callee is allowed to
temporarily overwrite these locations, as long as they are restored
once control returns. We will see variations of this later on: a
stronger, inductive integrity property that ensures that the caller's
private data is never overwritten, even during the caller's execution
(which leads to easier reasoning and more effective testing);
and a weaker, lazy integrity property that only ensures the preservation
of locations that subsequently affect the caller's observable actions 
(which allows for a more efficient enforcement mechanism).

@@

Trace integrity characterizes the trace corresponding to a single
callee.
Going back to our running example, such a trace would be, for
instance, the entire trace of {\tt h} from instruction 40 to
instruction 47. 
As an example of an integrity violation affecting this trace, \cref{fig:intex} shows what happens
if we change line 44 of the example program to $[\rsp - 5] \leftarrow r_4$.
Instead of storing data to the return value slot,
the changed version overwrites {\tt g}'s local variable {\tt z}.
The figure shows the modified trace, with snapshots of the initial
and final states colored to show the contour; high 
integrity memory is in \high, while low integrity is in \low.
At the end of the trace high integrity address 103 should match the initial
state, and does not. So trace integrity is violated.
\rb{There are some color name macros, but none for integrity yet? Use instead
  of hardcoded color names.}\apt{I don't understand what's going on with the colors.
  I thought we agreed to use the same coloring scheme throughout?}

We now lift this definition to an integrity property for a complete
execution trace. We do this by requiring that trace integrity holds
for every subtrace delimited by a call and its corresponding return.

\definition Given a trace \(\MPT\) and a call map \(\callmap\)
%and a function map \(\codemap\),
an \(n\)-{\em transition} is a pair of
machine-policy state pairs \(\mpstate{\mach_c}{\pol_c}\) and
\(\mpstate{\mach_e}{\pol_e}\) in the trace such that
\(\mach_c\) is a valid call with \(n\) arguments, and
\(\mpstate{\mach_c}{\pol_c}\) steps to \(\mpstate{\mach_e}{\pol_e}\).

\apt{or is in some function?? and why do we need either??}:
\sna{I think Andrew is right that we don't need to know the function
     here, or that the entry point is valid; that's part of WBCF.}
%and \(\mach_e\) is an entry point for some function \(f\):

  \[\begin{split}
    \mathit{Tr}_n ~ \MPT ~ \mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}
    \triangleq & \exists \obs, f.
    ~ \mpstate{\mach_c}{\pol_c} \stepstopol{\obs} \mpstate{\mach_e}{\pol_e} \\
    & \land \callmap(\mach_c(\PCname)) = n
      %\land \codemap(\mach_e(\PCname)) = f
  \end{split}\]

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\bcp{Hey, we should have an abstract \ldots}
\end{abstract}

@@

To phrase these two aspects of stack safety in terms of security properties
\leo{in terms of security properties or ``using security language''?}
requires an explicit attacker model: an clear understanding of who is being
protected and from what. Most existing stack protection mechanisms deal with
scenarios where a callee is influenced by external inputs and ends up
corrupting the stack in some way. To avoid getting into the details of
exactly how this happens in our formal properties, we use an even stronger
model: when stating the protections that a function can rely on when it
makes a call, we assume that the callee can execute arbitrary instructions
until it returns.

A key challenge in adapting the concepts of language-based security to this
setting is that stack safety is a ``nested'' property: not only does a given
caller need to be protected from its callee, but the callee also needs to be
protected from {\em its} callees, and so on. That is, the notion of
``attacker'' changes throughout the execution of the program.  We address
this issue using a dynamically changing notion of {\em accessibility} for
the different components of the machine state and adjusting the integrity
and confidentiality levels of different stack locations and registers at
every transition (call or return) between functions.

@@

\section{Threat Model}

@@

\bcp{More explanation needed, plus examples.}
A policy is a general model of an enforcement mechanism, consisting of a
set of policy states \(\pol \in \POLS\), a policy step function
\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
\rightharpoonup \POLS\), and a initializer \(\pinit : \MACHS \rightharpoonup
\POLS\). The initializer establishes the initial policy state for a given
machine state, but is partial to reflect that some policies are only defined
on states that are well-formed in some way.

We write $\MPS$ for the set of all pairs of machine states and policy
states.
%
\[\MPS = \MACHS \times \POLS\]
%
We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.

\leo{TODO: Forward pointer to policies right?}\bcp{Yes.  But also we can say
in more detail that the intention is to capture any kind of dynamic
enforcement mechanism and give as many names and citations as we can.}

@@

In this section we describe our machine model. Our definition of stack
safety is generic and this section discusses an abstract interface to
our machine model and some kind of static or dynamic, software- or
hardware-enforced security monitor. We will refer back to the running
example of the previous section for concreteness, and then,
in~\cref{sec:enforcement}, extend this model with a dynamic policy
enforcement mechanism based on
\citet{DBLP:conf/sp/RoesslerD18}.

@@

The basic building blocks of the machine are {\em values} and {\em
  addresses}. Both are drawn from some set of {\em words}
$\WORDS$, ranged over by $\word$.
%
Our machine states are composed of {\em components} $k$, addressed either by
words (memory addresses) or by by register names ($r$) drawn from some set
$\REGS$.
%
    \[\component \in \COMPONENTS = \WORDS + \REGS \]
%
The register names are assumed to include two special purpose registers: the
program counter {\PCname} and the stack pointer \SP.

@@

The step function of the machine takes a machine state and produces a
pair of a machine state and an {\em observation}, which may be either
silent (written $\tau$) or a word of data that the machine
communicates to the outside world (e.g., by writing to a DMA
address):
%
\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
\[\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \]
%

The choice of observations is an interesting one. A first approach
would be to consider any change to the registers or memory as observable.
However, as some parts of the stack will be considered ``private''
this choice won't work. A relaxation of this to consider all changes
to ``non-private'' data as observable could work in principle, but
does not allow for an interesting weaker-but-more-efficient enforcement
mechanism. We will come back to this point in \Cref{sec:lazy}. On the
other hand, we could use a particular location as an abstract output
of our system (e.g., through DMA). This is the approach we chose for
our running example to keep things simple. For that, $\WORDS$ is the
set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
\cup \{\PCname\}$, and $\SP = r_2$. Observations are the writes to the
special $\rout$ location: if during an execution step a word $w$ is
written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.

@@

In our model, the machine step function is total -- a program does not halt
on its own, but may instead enter a silent terminal loop. Runs are only cut
short by the policy step function. This means that software-only policies that
can catch errors must be modeled as reaching an error state, not outright
halting.

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by \(\stepstopol{}\). Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final state is
a \(\tau\):
\rb{This $\hookrightarrow_P$ is often enough written just $\hookrightarrow$.
Overload everywhere?}

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstoobs{\obs}
                 \mpstate{\mach_1}{\pol_1}\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow_P \MPT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P
                 \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
\end{minipage}

@@

\paragraph*{Until}
The ``Until'' operation \(\PUT(\Trace,f)\)
%\bcp{Wouldn't
%``While'' be more mnemonic?} \sna{Appealing, but it's more of a
%``while-not''? Trying out ``until''. In any case... moving away
%from doing joins means this needs to get some work to eliminate
%the trailing observation. Here's an attempt.}
takes a trace
\(\Trace : \mathit{trace} ~ \TRACEELEMS \times \OBSS\) and a predicate
on elements \(f \subset \TRACEELEMS\) and gives the prefix of
\(\Trace\) ending with the first element on which \(f\) holds.
Subtly, because the observation associated with a trace represents its
next step, when we take a prefix from a trace containing observations
we must replace the final observation with a \(\tau\).

  \begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f\ \traceelem\)}
             {\(\PUT(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
  \judgment{}
           {\(\PUT(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
              {\(\PUT(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
  \end{minipage}
  \end{center}
%
Note that if the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds,
\(\PUT\) is the identity function.

@@

\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
            {\(\PUT(f,\machT')
              = \machT\)}
            {\(\mach \hookrightarrow \machT | f\)}
\noindent
And similarly for traces with policies:

\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
            {\(\PUT(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
         {\(\mpstatename \hookrightarrow \MPT | f\)}

@@

We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
as far as an external observer is concerned, if the sequence of
non-silent observations of $\obsT_1$ is a prefix of those of
$\obsT_2$; that is we operate up to deletion of \(\tau\) observations,
coinductively:

@@

\leo{I still don't know how to do this transition. Maybe contrast
with the ``perhaps surprisingly'' later?} Note that
similarity of observation traces could alternatively be defined coinductively,
just like the prefix relation, but by dropping the assymetric rules that allow
for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$)

In addition, perhaps surprisingly, an infinite silent trace is a
prefix of (and equivalent to) any other trace. While this might seem
surprising at first, it makes sense in a timing-insensitive context:
an external observer looking at two machine runs cannot (computably)
distinguish between a machine that steps forever and a machine that
steps for a long time before producing some output.

%Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
%We then define {\em policy-sensitive observational equivalence}: a
%trace $\MPT \in \MPOTRACE$ is equivalent to a trace $\machT \in
%\MOTRACE$ if (1) \(\MPT\) is infinite and the observations of \(\MPT\)
%and \(\machT\) are equivalent, or (2) if the \(\MPT\) is finite (due to a
%policy fault) and its observations are a prefix of those of M.
%
%% Fixed by Sean
%%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
%\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
%            {\(\MPT \sim_P \machT\)}
%
%\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
%            {\(\MPT \sim_P \machT\)}

@@

In this section we will describe stack safety as a formal property. We
begin with our model of \leo{compiler-provided?} knowledge about the
program structure, and then dive into the formal definitions of local state
encapsulation and well-bracketed control flow that together constitute
stack safety.

@@

\paragraph*{Return Map}
Similarly, a {\em return map} identifies addresses that complete a
return sequence. Formally, it is simply a set of components; each one
is the final jump corresponding to a function:
%
\[\retmap \in \RETMAPS \subseteq \WORDS\]
%
In our running example, the return map identifies the $\jalr$
instructions 36 and 47:
\[ \retmap = \{36, 47\} \]

@@

We separately identify, for any call state, what it means for a state
to have correctly returned from a call. We consider the call to have
returned the first time control returns to the instruction following
the \jal with the stack restored. Formally this is a relation
between the states:

\[\begin{split}
    \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) =
    \nextPC(\mach_c(\PCname)) \land \\ & \mach_r(\SP) = \mach_c(\SP)
\end{split}\]

There are two such return points in our running example, both of which
are depicted in the figure: the state before the execution of
instruction 27 corresponding to the return from {\tt h} and the state
before instruction 8 corresponding to the return from {\tt g}.

@@

In the rest of this section we will implictly parameterize our
definitions by a function map $\codemap$, a call map $\callmap$, a return
map $\retmap$ and an entry map $\entmap$.

@@

When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) and
\(\mach_c \stepstoobs{\obs} \mach_e\) for some $n$, $\obs$, and
$\mach_e$, we can construct a contour based on the bounds of the
callee's stack frame.
%
The private state of the callee is the entire area of the memory from
stack base to stack pointer, with the exception of the $n$ locations
corresponding to the arguments and result locations of the
call. Private components are marked high confidentiality and high
integrity: the callee should be able to neither read, nor write them.
%
Everything above the stack pointer at the point of the call is
``uninitialized'' memory that can be written into but should not be
read. To protect the callee itself from such reads, these components
are marked high confidentiality and low integrity.
%
The same is true for the designated output location: it can be
freely written into but should not be read.
%
Special care is needed for the code memory corresponding to the
callee: they are marked low confidentiality if they belong to the
callee and high confidentiality otherwise; regardless they are always
high integrity.
%
Finally, registers and the frame of arguments to the callee are free
for the callee to use, and are therefore tagged with low integrity and
low confidentiality.

 \[\mathit{Cof}(\mach,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and for some } \funid, \codemap(\component) = \funid \\
    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \geq \mach(\SP)
                          \text{ or if } \component = \rout\\
    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \leq \mach(\SP) \negate  n \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]

Going back to our running example, recall the state of the machine at
the call from {\tt main} to {\tt g}, annotated with their
corresponding labels (not shown: the labels of the instructions
corresponding to {\tt g} that are $(\LOWSEC, \HIGHINT)$ instead of
$(\HIGHSEC, \HIGHINT)$):
\vspace*{0.2em}
% At Call to G
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{$(\HIGHSEC,\HIGHINT)$}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{\emptyoutc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{$(\HIGHSEC, \HIGHINT)$}}]%
\memory{2}{\fcallc}[{\makebox[0pt]{$(\LOWSEC, \LOWINT)$}}]%
\memory{6}{\unusedc}[{\makebox[0pt]{$(\HIGHSEC, \LOWINT)$}}]
~$\cdots$
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\MemoryLabel{-38.5em}{0.75em}{0}
\MemoryLabel{-34.5em}{0.75em}{0}
\\
\end{center}

@@

\paragraph*{Stack Integrity}

@@

Let $\contour$ be a contour and $\MPT$ be a (potentially infinite)
$\MPOTRACE$. We will say $\MPT$ satisfies {\em trace integrity} with
respect to $\contour$ if its final state (if any) agrees with its
initial one on all components marked $\HIGHSEC$ in $\contour$:
%
\[\begin{split}
  \mathit{int} ~ \contour ~ \MPT \triangleq &
    \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
  & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
  & \forall \component. ~ \contour(\component) = \HIGHSEC \Rightarrow
    \mach_e(\component) = \mach_r(\component) \\
\end{split}\]
%\leo{TODO: DISCUSS: Figure out notation for first/last machine state in a trace}
%\sna{We could do it like this, with finiteness implicit in the existence of
%     a final state.}+1

Intuitively, the caller expects to find their private data untouched
after the callee returns. That means that if the callee never returns,
integrity is trivially satisfied. Similarly, the callee is allowed to
temporarily overwrite these locations, as long as they are restored
once control returns. We will see variations of this later on: a
stronger, inductive integrity property that ensures that the caller's
private data is never overwritten, even during the caller's execution
(a variation that leads to easier reasoning and more effective testing);
and a weaker, lazy integrity property that only ensures that the
locations that a caller accesses are preserved (a variation that allows
for a more efficient enforcement mechanism).

@@

For an example see \cref{fig:intex}, representing the unsafe substitution
\[
  44: ~ \sw ~ \rsp ~ r_4 ~ -5
\]
This instruction is in {\tt h}, and instead of storing data to the return,
it overwrites {\tt g}'s variable, {\tt z}. At the top of the figure high
integrity memory is in red, while low integrity is in green; at the end of the
trace address 103 should match the initial state, and does not. So trace
integrity is violated.
\rb{There are some color name macros, but none for integrity yet? Use instead
of hardcoded color names.}

Trace integrity characterizes the trace corresponding to a single
callee. Going back to our running example, such a trace would be, for
instance, the entire trace of {\tt h} from instruction 40 to
instruction 47. This trace is repeated in \Cref{fig:intex}, with the initial
and final states colored to show the contour; high confidentiality
components in magenta, low in teal. At the end of the trace addresses 103
and 108 have both changed; since 103 was high confidentiality, the trace
does not enjoy trace integrity.

We now lift this definition to an integrity property for a complete
execution trace. We do this by requiring that trace integrity holds
for every transition between a caller and a callee for the entirety of
the trace of the callee.

\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
a function map $\codemap$, an $n$-{\em transition} is a pair of
machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
in the trace such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

\definition{Stack Integrity} Let $\callmap$ be a call map and
$\codemap$ a function map. We say that a system satisfies {\em stack
  integrity} with respect to $\callmap$ and $\codemap$ if, for every
initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
(if any) satisfies trace integrity with respect to the contour at the
call:
%
$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$

\paragraph*{Stack Confidentiality}

The second component of local state encapsulation is a confidentiality
property: just like integrity ensures that a caller's data is never
overwritten by a callee, confidentiality ensures that it never
influences its observable behavior. We will formally define
confidentiality in two similar steps: a trace version and a system
version.
